---
title: "InClassChallenge6"
author: "Alex Berry"
date: "2025-03-27"
output:
  html_document:
    toc: true
  md_document:
    variant: gfm
  word_document:
    toc: true
  pdf_document: null
---

# QUESTION 1
The main point of writing your own functions is to avoid copy and paste errors when you need to perform the same code on different data. Functions allow you to simplify your code, allowing for better data management and reproducibility. Iterations are also helpful to reduce copy and paste errors when you need to repeat code multiple times, ultimately expediting repetitive work. What both functions and iterations accomplish is the reduction of copy and paste errors, and allows for consistent, and reusable code. 

# QUESTION 2
```{r}
# Writing a function
function_name <- function(factor1, factor2) {
  result <- factor1 + factor2
  return(result)
}
```

```{r}
# Example
add_numbers <- function(a, b) {
  sum <- a + b
  return(sum)
}

add_numbers(12, 15)
```

```{r}
# Writing a For Loop
for (i in 1:5) {
 print(i)
}
```

```{r}
# Example
for (i in 1:5) {
  print(i)
}
```

The code will be written in R Script or R Markdown. The output will be in the console. A function will grooup code to do a specific task once defined. A for loop will repeat a block of code a specific amount of times so you can perform the same function on a sequence of values.

# QUESTION 3
```{r}
cities.data <- read.csv("Cities.csv")
```

# QUESTION 4
```{r}
haversine_distance <- function(lat1, lon1, lat2, lon2) {
  rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a)) 
  earth_radius <- 6378137
  distance_km <- (earth_radius * c)/1000
  return(distance_km)
}
```

# QUESTION 5
```{r}
library(tidyverse)
auburn_lat <- subset(cities.data, city == "Auburn" & state_id == "AL") %>%
  pull(lat)
auburn_long <- subset(cities.data, city == "Auburn" & state_id == "AL") %>%
  pull(long)
nyc_lat <- subset(cities.data, city == "New York" & state_id == "NY") %>%
  pull(lat)
nyc_long <- subset(cities.data, city == "New York" & state_id == "NY") %>%
  pull(long)

haversine_distance(auburn_lat, auburn_long, nyc_lat, nyc_long)
```

# QUESTION 6
```{r}
head(cities.data, 10)

auburn_lat <- subset(cities.data, city == "Auburn" & state_id == "AL") %>%
  pull(lat)
auburn_long <- subset(cities.data, city == "Auburn" & state_id == "AL") %>%
  pull(long)

for (i in 1:nrow(cities.data)) {
    lat_city <- cities.data$lat[i]
    lon_city <- cities.data$long[i]
    distances <- haversine_distance(auburn_lat, auburn_long, lat_city, lon_city)
    print(distances)
}

cities.data$Distance_from_Auburn_km <- distances

head(cities.data, 10)

# Bonus
auburn_lat <- subset(cities.data, city == "Auburn" & state_id == "AL") %>%
  pull(lat)
auburn_long <- subset(cities.data, city == "Auburn" & state_id == "AL") %>%
  pull(long)

distance_df <- data.frame(city1 = character(0), city2 = character(0), distance_km = numeric(0))

for (i in 1:nrow(cities.data)) {
  if (cities.data$city[i] != "Auburn" | cities.data$state_id[i] != "AL") {
    lat_city <- cities.data$lat[i]
    lon_city <- cities.data$long[i]
    distance <- haversine_distance(auburn_lat, auburn_long, lat_city, lon_city)
    new_row <- data.frame(city1 = cities.data$city[i], city2 = "Auburn", distance_km = distance)
    distance_df <- rbind(distance_df, new_row)
  }
}

head(distance_df, 10)
```

# QUESTION 7
[Link to GitHub](https://github.com/alexberry8/In-Class-Coding-Challenges)





